%\vspace{-4.0\baselineskip}
%\vspace{4 cm}
%\begin{flushright}
%    {\LARGE \textbf{ABSTRACT}}
%\end{flushright}

%\vspace{-0.5\baselineskip}
%\noindent\rule{\linewidth}{2pt}
\newgeometry{tmargin=2.5cm, lmargin=38mm, rmargin=25mm, bmargin=25mm}
\thispagestyle{empty}
%\vspace*{0.1cm}
\begin{center}
{\underline{\bf\Large ABSTRACT }}\\
\par\vspace{5mm}
\end{center}

The exponential growth of textual data has introduced significant challenges in effectively retrieving and ranking relevant documents for information-intensive tasks. Traditional information retrieval (IR) systems often fall short in capturing semantic nuances and providing contextually accurate rankings. This thesis proposes a novel approach to document ranking using Retrieval-Augmented Generation (RAG)\cite{rag} enhanced with Hypothetical Document Embedding (HyDE)\cite{hyde}. 

The proposed framework integrates dense retrieval techniques with generative language models to achieve superior ranking accuracy. Initially, candidate documents are retrieved using dense embeddings based on transformer models. To address ambiguity and incomplete queries, the HyDE component generates hypothetical answers, which are encoded and utilized to refine the ranking process. A generative model further evaluates the contextual relevance of the candidate documents and adjusts the ranking accordingly. The final system leverages a hybrid pipeline that combines retrieval, generation, and reranking to address key challenges in IR.

Comprehensive experiments on benchmark datasets demonstrate that the proposed methodology significantly outperforms traditional ranking techniques in terms of relevance metrics such as Accuracy, Mean Reciprocal Rank (MRR). By advancing the integration of retrieval and generation, this work provides a robust framework for next-generation IR systems and sets a foundation for future research in knowledge-intensive tasks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%