\chapter{Conclusions and Future Scope }
\label{C5} %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
% \label{C5} %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}
The proposed methodology of fine-tuning pre-trained LLMs for cardiovascular diseases has been successfully implemented. The fine-tuned models, Gemini\_PMC and GPT2\_PMC, have shown improved performance on the medical dataset, outperforming the base models. The results demonstrate the effectiveness of fine-tuning large language models for medical domain-specific tasks, highlighting the potential of LLMs in healthcare applications. 
\section{Future Scope}
\begin{enumerate}
    \item In future a better curated set of data should be used for fine-tuning the models. As of today there is a lack of resource available for the specific field of cardiovascular diseases.
    \item Other PEFT techniques such as LoRA, and many other adapter based methods can be used for faster fine-tuning of the models.
    \item The models can be further evaluated on a larger and more diverse set of medical datasets to assess their capabilities. An approach that may be used is by involving medical professionals to rate responses to the questions set by them.
    \item The proposed methodology can be extended to other medical domains, such as oncology, neurology, and radiology, to develop specialized LLMs for various healthcare applications.
    \item The future work can also include the development of multimodal LLMs that combine text, image, and other data modalities to enhance the performance of medical models. 
    \item Involving LLMs into healthcare has a huge risk associated with it. To mitigate this risk, the models can be further tuned on an extensive set of ethical guidelines. This won't eliminate the risk but will keep the model in check to produce reponses that can cause potential damage to health without a medical expert's supervision.
\end{enumerate}


% \begin{enumerate}[label=(\roman*)]
%     \item More detailed high-resolution thermal images can be implemented for better enhancement of important features.
%     \item Other updated deep-learning algorithms can be implemented for better flaws identification.
%     \item For improvement of the performance of the fusion algorithm with optimization techniques, other optimizers can be utilized.
% \end{enumerate}